Large-scale hybrid ad hoc network for mobile platforms: Challenges and Experiences

### 2019-02-11 paper.tex
First complete draft, needs shortening and reviewing

\section{Introduction}
Mobile devices with programmable platforms such as android and iOS
have steadily grown over the last decade, surpassing the 2 billion
mark \footnote{https://www.statista.com/statistics/330695/number-of-smartphone-users-worldwide/}. MANETs
have been widely studied given their decentralized nature and
potential for new applications
\cite{loo-manet-2011,perkins-ad-hoc-2001}. Most of the prior work on
p2p networks has focused on analytical and simulation-based study of
MANET behavior
\cite{zhang-topology-manet-2015,marti-misbehavior-manet-2000,mauve-pos-routing-manet-2001}. Given
the outstanding practical challenges in deployoing a large number of
physical nodes, real-world implementations have been limited and have
not reached mass scale \cite{kiess-manet-impl-2007}. However, with the
growth of smartphones, large-scale real-world implementations may
become feasible. This paper describes a real-world implementation of a
p2p delay-tolerant network, called Mesh Network Alerts (MNA), for
relaying severe weather information to millions of mobile device users
as part of the Weather Channel app \cite{mna} on both android and iOS
platforms.

@@

The above needs are well-recognized in the industry as well as
academia with several ambitious attempts to address them, e.g., Google
Loon project\footnote{https://loon.co} and Facebook
Aquilla\footnote{https://en.wikipedia.org/wiki/Facebook\_Aquila},
though with limited impact. Leveraging user mobile devices as peer
nodes for a large-scale deployment has been another theme in the prior
works, e.g., the Serval project
\cite{gardner-stephen-serval-2011}. Serval mesh enables p2p
communication over on-device WiFi radio, but requires root access to
the device via jailbreaking. Although significant leassons have been
learned through these attempts, a mass-scale p2p network for such
applications remains elusive.

@@

continue to discover peer devices to receive and formward

@@

connection latencies and wireless transmission are specific to device

@@

As user devices with MNA advertise on unsecure wireless protocols, it

@@

To address (or bypass) the challenges identified above, this paper
proposes c2p2p -- a hybrid architecture that leverages a central
service as the sole source of trusted information as depicted in
Figure~\ref{fig:arch}. An application-specific central service, e.g.,
weather.com backend, originates new information as a push message and
assigns it a unique identifier. Next, the central service produces a
digital signature using its private key and appends it to the message
payload. Finally, header parameters specifying TTL (duration after
which the message expires), peer identifier of the central service,
destination peer identifier list (for unicast) or a special identifier
(for broadcast), and the time of origin are added to the message and
the message is sent to a push notification service for distribution.

Mobile applications integrate MNA as an SDK including the public key
of the corresponding central service. Peers having Internet
connectivity receive the push messages and verify the digital
signature to protect against spread of misinformation.  If the message
has not expired and is destined to other peers, then the receiving
peers start forwarding the message to other peers. Depending on the
protocol being used, such forwarding happens as a unicast or
broadcast.

@@

The following assesses the c2p2p architecture relative to the
challenges identified above along with known limitations. Further, to
realize this architecture on mobile devices, we evaluated a variety of
protocols on both Android and iOS. As described later, WiFi DNS on
Android, Bluetooth Nearby on Android, and Blutooth low energy (BTLE)
on both Android and iOS are the only protocols that were found
effective. Actual techniques for discovery, advertisement, and
connection vary across these protocols and are described next.
%
\subsection{Operating system restrictions}
\label{sec:os}
MNA works around the background activity and user permission
restrictions via innovative techniques, without resorting to hacks
that may violate user security or App-store guidelines.
Tables~\ref{tab:android} and~\ref{tab:ios} summarize the various
protocols we have experimented with using WiFi and Bluetooth
interfaces, along with key findings as described here.  The protocols
actually deployed are in bold.
%
%
\begin{table}[H]
\caption{Android protocols and key findings}
\label{tab:android}
\centering
\begin{tabular}{|c|l|r|r|r|c|l|}
\hline
& \multicolumn{1}{c|}{Protocol} & \multicolumn{1}{c|}{Throughput} & \multicolumn{1}{c|}{Battery} & \multicolumn{1}{c|}{Range} & \multicolumn{1}{c|}{iOS} & \multicolumn{1}{c|}{Issues}\\
&                               & \multicolumn{1}{c|}{Avg. kbps}  & \multicolumn{1}{c|}{per hour} & \multicolumn{1}{c|}{LoS ft} &                  &                            \\
\hline
\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{WiFi}}} & \textbf{DNS} & 0.2 & 2\% & 600 & No & \\
& WiDi & 2000 & 3\% & 600 & No & Needs DNS\\
& Hotspot & 2000 & 5\% & 600 & No & Permission\\
\hline
\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{BT}}} & \textbf{Classic} & 50 & 2\% & 600 & No & \\
& \textbf{Nearby} & 50 & 2\% & 600 & No & \\
& \textbf{BTLE} & 50 & 2\% & 600 & Yes & \\
\hline
\end{tabular}
\end{table}
%
%
\begin{table}[H]
\caption{iOS protocols and key findings}
\label{tab:ios}
\centering
\begin{tabular}{|c|l|r|r|r|c|l|}
\hline
& \multicolumn{1}{c|}{Protocol} & \multicolumn{1}{c|}{Throughput} & \multicolumn{1}{c|}{Battery} & \multicolumn{1}{c|}{Range} & \multicolumn{1}{c|}{Android} & \multicolumn{1}{c|}{Issues}\\
&                               & \multicolumn{1}{c|}{Avg. kbps}  & \multicolumn{1}{c|}{per hour} & \multicolumn{1}{c|}{LoS ft} &                  &                            \\
\hline
\parbox[t]{2mm}{\multirow{2}{*}{\rotatebox[origin=c]{90}{WiFi}}} & Bonjour & 0.2 & 5\% & 200 & No & Battery\\
& MPC & 2000 & 5\% & 200 & No & Battery\\
\hline
& \textbf{BTLE} & 50 & 1\% & 800 & Yes & \\
\hline
\end{tabular}
\end{table}
%

@@

%
\subsection{Power constraints}
\label{sec:power}
%
Due to indefinite forground services and continuous discovery and
advertisement, this is primarily an issue on Android. Our approach
here is two-pronged. Firstly, via extensive experiments on a large
number of devices, we fine-tune the algorithms governing the intervals
at which discovery and advertisements occur. In a nutshell, receiving
new information during a period causes MNA to be more aggressive in
discovery and advertisement. Similarly, lack of new information for a
period makes the device less aggressive. Secondly, we allow the
centrasl service to broadcast``wake up'' messages ahead of an
anticipated severe weather event.  When devices receive such messages,
they schedule themselves to remain aggressive during the specified
window of time.  Outside of this window, the device can afford to have
long sleep cycles and conserve power. With these techniques, our
testing shows less than 2\% battery consumption per hour on most
device models.
%
\subsection{Testing p2p networks}
\label{sec:texting}
%
We developed novel test automation tools and processes that control
multiple devices from a single test station and follow prescribed
steps to generate, send, and receives messages to play out a test
scenario.  The framework allows automated analysis of the observations
to determine the test result. This capability was instrumental in
uncovering bugs at a fast pace with tens of physical devices emloyed
for automated testing.  Further, the automation framework is general
and can be expressly applied to test other apps in this fashion.

Figure~\ref{fig:test_arch} shows the overall test automnation
architecture. A tester specifies high-level test cases and expected
results in a platform and device agnostic fashion. Mobile devices run
a UI proxy for executing platform and device-specific UI command while
a central laptop terminal acts as the overall orchestrator and
translates high-level test cases and distributes them to
devices. Since the orchestrator needs a wired access to the devices
and the devices cannot be programmed to move (unlike drones and bots),
testing of range and mobility patterns needs to done manually, which
is a known limitation.

@@

Peers cannot originate messages however as then a malicious peer can
join the network, exchange keys with other peers, start generating,
signing, and spreading fake information. This has been a longstanding
problem for open decentralized systems. Requiring all peers to
register their identity with a certificate authority and reporting the
originators of fake information may discourage malicious behavior.
Clearly, this is even harder to solve than the spread of fake
information over the Internet because ground truth is hard to find and
malicious peers can report genuine sources of information as well.
%
\section{Algorithms}
\label{sec:algo}
%
\subsection{WiFi DNS}
\label{sec:wifi}
%
DNS service discovery is a widely supported protocol
\cite{cheshire-dns-sd-2013}. On Android, WiFi DNS improvises on this
standard such that the exchange of information happens without making
network connections at all. This is achieved by splitting the message
payloads into small enough chunks and stuffing them into service
advertisements as txt records and broadcast over multiple
advertisements in a quick succession. Since there are no connections
being made, the messages can be exchanged without prompting the user
to grant permissions for each connection. Since no peer connections
are needed and advertisements and discovery are broadcast, WiFi DNS is
a broadcast protocol. On iOS, Bonjour protocol is built on DNS service
discovery but is power-hungry and does not interoperate with WiFi DNS
on Android. Algorithm~\ref{algo:dns} outlines the peer process for
WiFi DNS. Power conservation via time intervals is achieved by
manupulating $T_{ad}$ and $T_{disc}$ based on traffic.

Messages from multiple applications are forwarded on the same network,
but a peer can validate signatures only for the messages targeting an
app it hosts. Peers simply forward messages not targeting one of their
apps. Key benefit of this design is that the density of peers is
available to all apps, making c2p2p scenarios viable even when an
individual app does not have a sufficient density of users.  However,
this also allows a malicioous jammer to generate traffic for a fake
app and force the peers to forwad it (as they do not match any of the
apps) until the messages expire.  Although the fake information will
never be delivered to the apps, this issue can affect MNA throughput
for legitimate apps. A possible solution would be to maintain a
central registry of $a_{id}$ and distribute it to all peers across
apps.

\begin{algorithm}[htb!]
\label{algo:dns}
\DontPrintSemicolon % Some LaTeX compilers require you to use \dontprintsemicolon instead
\SetAlgoLined
\KwIn{Peer id $p_{id}$}
\KwIn{App id $a_{id}$, max advert/discovery intervals $max_{ad}$, $max_{disc}$}
\KwIn{min advert/discovery intervals $min_{ad}$, $min_{disc}$}
 $S_{name} \gets$ \textsf{'MNA'}\;
 $\mathbb{M} \gets \varnothing$\;
 $T_{ad} \gets min_{ad}$\;
 $T_{disc} \gets min_{disc}$\;
 $New \gets 0$
 \ForEach{timer expiry $T_{ad}$}{
   $\mathbb{M} \gets \mathbb{M} -$ expired messages\;
   $\mathbb{M}_{sort} \gets$ $\mathbb{M}$ sorted by least advertised first\;
   $B \gets$ $\mathbb{M}_{sort}$ chunked by MTU size\;
   dnsAdvertise ($S_{name}$, $p_{id}$, $B$)\;
   \If{$New = 0$}{
     $T_{ad} \gets min(T_{ad} \times 2, max_{ad})$\;
   }
 }
 \ForEach{timer expiry $T_{disc}$}{
   dnsDiscover ($S_{name}$, $p_{id}$)\;
   \If{$New = 0$}{
     $T_{disc} \gets min(T_{disc} \times 2, max_{disc})$\;
   }
   $New \gets 0$\;
 }
 \ForEach{discovered service callback $S_{disc}$}{
   $B \gets S_{disc}.B$\;
   $M_{part} \gets msg(B)$\;
   $M \gets M + M_{part}$\; 
   \If{$M_{part}$ is not a duplicate}{
        $T_{ad} \gets max(\frac{T_{ad}}{2}, min_{ad})$\;
        $T_{disc} \gets max(\frac{T_{disc}}{2}, min_{disc})$\;
        $New \gets 1$\;
   }
   \If{$M_{part}$ was the last chunk of $M$}{
     \If{$M$ is unexpired}{
         \If{$M.app = a_{id}$ $\wedge$ $M$ signature is valid}{
             Notify app with $M$\;
         }
         $\mathbb{M} \gets \mathbb{M} \cup M$\;
     }
   }
 }
 \tcc{For testing only}
 \ForEach{app request to send $M$}{
   \If{$M$ has not expired}{
     $\mathbb{M} \gets \mathbb{M} \cup M$\;
   }
 }
 \caption{WiFi DNS peer algorithm}
\end{algorithm}
%
\subsection{BTLE}
\label{sec:btle}
%
The only protocol that stays reliably active in
background without violating iOS developer guidelines is BTLE. Unlike
Android, iOS applications do not need to request discovery and
advertisements contunually. Once a BTLE Peripheral (role that has
information) is advertised and a Central (role that receives
information) requests discovery, the app is allowed to go to
sleep. When a matching central or a peripheral device is found, iOS
wakes up the application app to handle such an event, even when the
application is in the background. MNA builds on this and leverages
acynchronous dispatch queues to turn each peer into a Central as well
as a Peripheral simultaneously. This architecture allows bidirectional
data transfers without concurrency issues.

@@

\section{Experimental Results}
\label{sec:eval}
%

%
\subsection{Pilot test results}
\label{sec:romania}
%
\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/romania_latency}}
\caption{Latency results from a 25-user pilot}
\label{fig:romania_lat}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/romania_activity}}
\caption{MNA activity during the pilot}
\label{fig:romania_act}
\end{figure}

About 13\% of the messages were received at least once (depends on
density of users in the building). Users actively used the devices for
~5\% of the time p2p NSD achieved lower end-to-end latencies than BT
Classic, perhaps due to its broadcast model and wide support across
devices. Packet sizes did not affect single-hop latencies on p2p DNS,
perhaps because most of the messages fit within single advert. Packet
sizes did not affect single-hop latencies on BT Classic, perhaps
because Bluetooth has a much higher bandwidth 2 out of 25 devices have
their clocks set wrong (confusing latency results). More than 10\% of
the messages were received in the last 5 min of message expiry,
implying messages were discarded due to expirty, even when they did
not reach all devices

%
\subsection{Automated test results}
\label{sec:automated}
%
\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/variety_e2e_latency}}
\caption{Comparison of WiFi and Bluetooth interfaces in terms of
  end-to-end latency}
\label{fig:variety_e2e}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/variety_hops}}
\caption{Comparison of WiFi and Bluetooth interfaces in terms of
  network topology}
\label{fig:variety_hops}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/e2e_latency}}
\caption{Comparison of WiFi and Bluetooth Nearby interfaces in terms
  of end-to-end latency}
\label{fig:e2e}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/duplicates}}
\caption{Duplicity due to flooding}
\label{fig:dup}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/hops}}
\caption{Average hop count}
\label{fig:hop}
\end{figure}

%
\section{Related Work}
\label{sec:related}
%
Apart from the state-of-the-art cited earlier, there have been other
notable attempts at practical large-scale deployments and we describe
them here. Before mid-2000’s, most of the research on MANETs was based
on Department of Defense requirements, until commodity multi-hop ad
hoc networks began to be considered \cite{bruno-mesh-2005}. However,
it took another decade before wireless mesh networking was used
commercially to enable smartphones to connect via Bluetooth and WiFi
in a popular application called FireChat \cite{firechat}. The success
of FireChat, partially due to the news coverage of its use in
political situations in which governments restricted access to the
Internet, has led to many alternatives in the past few years. An
up-to-date list of such applications is available
here \footnote{https://alternativeto.net/software/firechat-by-open-garden/}
\section{Conclusions and Future Work}
\label{sec:conclude}

\section*{Acknowledgment}
This research was sponsored by the U.S. Army Research Laboratory and
the U.K. Ministry of Defence under Agree- ment Number
W911NF-16-3-0001. The views and conclusions contained in this document
are those of the authors and should not be interpreted as representing
the official policies, either expressed or implied, of the U.S. Army
Research Laboratory, the U.S. Government, the U.K. Ministry of Defence
or the U.K. Government. The U.S. and U.K. Governments are au- thorized
to reproduce and distribute reprints for Government purposes
notwithstanding any copyright notation hereon.
\bibliographystyle{abbrv} \bibliography{smartcomp19}

>>>>

\section{Introduction}
Mobile devices with programmable platforms such as android and iOS
have steadily grown over the last decade, surpassing the 2 billion
mark \footnote{https://www.statista.com/statistics/330695/number-of-smartphone-users-worldwide/}. MANETs
have been widely studied given their decentralized nature and
potential for new applications
\cite{loo-manet-2011,perkins-ad-hoc-2001}. Most of the prior work on
p2p networks has focused on analytical and simulation-based study of
MANET behavior
\cite{zhang-topology-manet-2015,marti-misbehavior-manet-2000,mauve-pos-routing-manet-2001}. Given
the outstanding practical challenges in deploying a large number of
physical nodes, real-world implementations have been limited and have
not reached mass scale \cite{kiess-manet-impl-2007}. However, with the
growth of smartphones, large-scale real-world implementations may
become feasible. This paper describes a real-world implementation of a
p2p delay-tolerant network, called Mesh Network Alerts (MNA), for
relaying severe weather information to millions of mobile device users
as part of the Weather Channel app \cite{mna} on both android and iOS
platforms.

@@

The above needs are well-recognized in the industry as well as
academia with several ambitious attempts to address them, e.g., Google
Loon project\footnote{https://loon.co} and Facebook
Aquilla\footnote{https://en.wikipedia.org/wiki/Facebook\_Aquila},
though with limited impact. Leveraging user mobile devices as peer
nodes for a large-scale deployment has been another theme in the prior
works, e.g., the Serval project
\cite{gardner-stephen-serval-2011}. Serval mesh enables p2p
communication over on-device WiFi radio, but requires root access to
the device via jail-breaking. Although significant lessons have been
learned through these attempts, a mass-scale p2p network for such
applications remains elusive.

@@

continue to discover peer devices to receive and forward

@@

connection latency and wireless transmission are specific to device

@@

As user devices with MNA advertise on unsecured wireless protocols, it

@@

To address (or bypass) the challenges identified above, this paper
proposes c2p2p -- a hybrid architecture that leverages a central
service as the sole source of trusted information as depicted in
Figure~\ref{fig:arch}. An application-specific central service, e.g.,
weather.com backend, originates new information as a push message and
assigns it a unique identifier. Next, the central service produces a
digital signature using its private key and appends it to the message
payload. Finally, header parameters specifying TTL (duration after
which the message expires), peer identifier of the central service,
destination peer identifier list (for uni-cast) or a special identifier
(for broadcast), and the time of origin are added to the message and
the message is sent to a push notification service for distribution.

Mobile applications integrate MNA as an SDK including the public key
of the corresponding central service. Peers having Internet
connectivity receive the push messages and verify the digital
signature to protect against spread of misinformation.  If the message
has not expired and is destined to other peers, then the receiving
peers start forwarding the message to other peers. Depending on the
protocol being used, such forwarding happens as a uni-cast or
broadcast.

@@

The following assesses the c2p2p architecture relative to the
challenges identified above along with known limitations. Further, to
realize this architecture on mobile devices, we evaluated a variety of
protocols on both Android and iOS. As described later, WiFi DNS on
Android, Bluetooth Nearby on Android, and Bluetooth low energy (BTLE)
on both Android and iOS are the only protocols that were found
effective. Actual techniques for discovery, advertisement, and
connection vary across these protocols and are described next.
%
\subsection{Operating system restrictions}
\label{sec:os}
MNA works around the background activity and user permission
restrictions via innovative techniques, without resorting to hacks
that may violate user security or App-store guidelines.
Tables~\ref{tab:android} and~\ref{tab:ios} summarize the various
protocols we have experimented with using WiFi and Bluetooth
interfaces, along with key findings as described here.  The protocols
actually deployed are in bold.
%
%
\begin{table}[H]
\caption{Android protocols and key findings}
\label{tab:android}
\centering
\begin{tabular}{|c|l|r|r|r|c|l|}
\hline
& \multicolumn{1}{c|}{Protocol} & \multicolumn{1}{c|}{Throughput} & \multicolumn{1}{c|}{Battery} & \multicolumn{1}{c|}{Range} & \multicolumn{1}{c|}{iOS} & \multicolumn{1}{c|}{Issues}\\
&                               & \multicolumn{1}{c|}{Avg. kbps}  & \multicolumn{1}{c|}{per hour} & \multicolumn{1}{c|}{LoS ft} &                  &                            \\
\hline
\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{WiFi}}} & \textbf{DNS} & 0.2 & 2\% & 600 & No & \\
& WiDi & 2000 & 3\% & 600 & No & Needs DNS\\
& Hotspot & 2000 & 5\% & 600 & No & Permission\\
\hline
\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{BT}}} & \textbf{Classic} & 50 & 2\% & 600 & No & \\
& \textbf{Nearby} & 50 & 2\% & 600 & No & \\
& \textbf{BTLE} & 50 & 2\% & 600 & Yes & \\
\hline
\end{tabular}
\end{table}
%
%
\begin{table}[H]
\caption{iOS protocols and key findings}
\label{tab:ios}
\centering
\begin{tabular}{|c|l|r|r|r|c|l|}
\hline
& \multicolumn{1}{c|}{Protocol} & \multicolumn{1}{c|}{Throughput} & \multicolumn{1}{c|}{Battery} & \multicolumn{1}{c|}{Range} & \multicolumn{1}{c|}{Android} & \multicolumn{1}{c|}{Issues}\\
&                               & \multicolumn{1}{c|}{Avg. kbps}  & \multicolumn{1}{c|}{per hour} & \multicolumn{1}{c|}{LoS ft} &                  &                            \\
\hline
\parbox[t]{2mm}{\multirow{2}{*}{\rotatebox[origin=c]{90}{WiFi}}} & Bonjour & 0.2 & 5\% & 200 & No & Battery\\
& MPC & 2000 & 5\% & 200 & No & Battery\\
\hline
& \textbf{BTLE} & 50 & 1\% & 800 & Yes & \\
\hline
\end{tabular}
\end{table}
%

@@

%
\subsection{Power constraints}
\label{sec:power}
%
Due to indefinite foreground services and continuous discovery and
advertisement, this is primarily an issue on Android. Our approach
here is two-pronged. Firstly, via extensive experiments on a large
number of devices, we fine-tune the algorithms governing the intervals
at which discovery and advertisements occur. In a nutshell, receiving
new information during a period causes MNA to be more aggressive in
discovery and advertisement. Similarly, lack of new information for a
period makes the device less aggressive. Secondly, we allow the
central service to broadcast``wake up'' messages ahead of an
anticipated severe weather event.  When devices receive such messages,
they schedule themselves to remain aggressive during the specified
window of time.  Outside of this window, the device can afford to have
long sleep cycles and conserve power. With these techniques, our
testing shows less than 2\% battery consumption per hour on most
device models.
%
\subsection{Testing p2p networks}
\label{sec:texting}
%
We developed novel test automation tools and processes that control
multiple devices from a single test station and follow prescribed
steps to generate, send, and receives messages to play out a test
scenario.  The framework allows automated analysis of the observations
to determine the test result. This capability was instrumental in
uncovering bugs at a fast pace with tens of physical devices employed
for automated testing.  Further, the automation framework is general
and can be expressly applied to test other apps in this fashion.

Figure~\ref{fig:test_arch} shows the overall test automation
architecture. A tester specifies high-level test cases and expected
results in a platform and device agnostic fashion. Mobile devices run
a UI proxy for executing platform and device-specific UI command while
a central laptop terminal acts as the overall orchestrator and
translates high-level test cases and distributes them to
devices. Since the orchestrator needs a wired access to the devices
and the devices cannot be programmed to move (unlike drones and bots),
testing of range and mobility patterns needs to done manually, which
is a known limitation.

@@

Peers cannot originate messages however as then a malicious peer can
join the network, exchange keys with other peers, start generating,
signing, and spreading fake information. This has been a longstanding
problem for open decentralized systems. Requiring all peers to
register their identity with a certificate authority and reporting the
originators of fake information may discourage malicious behavior.
Clearly, this is even harder to solve than the spread of fake
information over the Internet because ground truth is hard to find and
malicious peers can report genuine sources of information as well.
%
\section{Algorithms}
\label{sec:algo}
%
\subsection{WiFi DNS}
\label{sec:wifi}
%
DNS service discovery is a widely supported protocol
\cite{cheshire-dns-sd-2013}. On Android, WiFi DNS improvises on this
standard such that the exchange of information happens without making
network connections at all. This is achieved by splitting the message
payloads into small enough chunks and stuffing them into service
advertisements as txt records and broadcast over multiple
advertisements in a quick succession. Since there are no connections
being made, the messages can be exchanged without prompting the user
to grant permissions for each connection. Since no peer connections
are needed and advertisements and discovery are broadcast, WiFi DNS is
a broadcast protocol. On iOS, Bonjour protocol is built on DNS service
discovery but is power-hungry and does not inter-operate with WiFi DNS
on Android. Algorithm~\ref{algo:dns} outlines the peer process for
WiFi DNS. Power conservation via time intervals is achieved by
manipulating $T_{ad}$ and $T_{disc}$ based on traffic.

Messages from multiple applications are forwarded on the same network,
but a peer can validate signatures only for the messages targeting an
app it hosts. Peers simply forward messages not targeting one of their
apps. Key benefit of this design is that the density of peers is
available to all apps, making c2p2p scenarios viable even when an
individual app does not have a sufficient density of users.  However,
this also allows a malicious jammer to generate traffic for a fake
app and force the peers to forward it (as they do not match any of the
apps) until the messages expire.  Although the fake information will
never be delivered to the apps, this issue can affect MNA throughput
for legitimate apps. A possible solution would be to maintain a
central registry of $a_{id}$ and distribute it to all peers across
apps.

\begin{algorithm}[htb!]
\label{algo:dns}
\DontPrintSemicolon % Some LaTeX compilers require you to use \dontprintsemicolon instead
\SetAlgoLined
\KwIn{Peer id $p_{id}$}
\KwIn{App id $a_{id}$, max advert/discovery intervals $max_{ad}$, $max_{disc}$}
\KwIn{min advert/discovery intervals $min_{ad}$, $min_{disc}$}
 $S_{name} \gets$ \textsf{'MNA'}\;
 $\mathbb{M} \gets \varnothing$\;
 $T_{ad} \gets min_{ad}$\;
 $T_{disc} \gets min_{disc}$\;
 $New \gets 0$
 \ForEach{timer expiry $T_{ad}$}{
   $\mathbb{M} \gets \mathbb{M} -$ expired messages\;
   $\mathbb{M}_{sort} \gets$ $\mathbb{M}$ sorted by least advertised first\;
   $B \gets$ $\mathbb{M}_{sort}$ chunked by MTU size\;
   dnsAdvertise ($S_{name}$, $p_{id}$, $B$)\;
   \If{$New = 0$}{
     $T_{ad} \gets min(T_{ad} \times 2, max_{ad})$\;
   }
 }
 \ForEach{timer expiry $T_{disc}$}{
   dnsDiscover ($S_{name}$, $p_{id}$)\;
   \If{$New = 0$}{
     $T_{disc} \gets min(T_{disc} \times 2, max_{disc})$\;
   }
   $New \gets 0$\;
 }
 \ForEach{discovered service callback $S_{disc}$}{
   $B \gets S_{disc}.B$\;
   $M_{part} \gets msg(B)$\;
   $M \gets M + M_{part}$\; 
   \If{$M_{part}$ is not a duplicate}{
        $T_{ad} \gets max(\frac{T_{ad}}{2}, min_{ad})$\;
        $T_{disc} \gets max(\frac{T_{disc}}{2}, min_{disc})$\;
        $New \gets 1$\;
   }
   \If{$M_{part}$ was the last chunk of $M$}{
     \If{$M$ is unexpired}{
         \If{$M.app = a_{id}$ $\wedge$ $M$ signature is valid}{
             Notify app with $M$\;
         }
         $\mathbb{M} \gets \mathbb{M} \cup M$\;
     }
   }
 }
 \tcc{For testing only}
 \ForEach{app request to send $M$}{
   \If{$M$ has not expired}{
     $\mathbb{M} \gets \mathbb{M} \cup M$\;
   }
 }
 \caption{WiFi DNS peer algorithm}
\end{algorithm}
%
\subsection{BTLE}
\label{sec:btle}
%
The only protocol that stays reliably active in
background without violating iOS developer guidelines is BTLE. Unlike
Android, iOS applications do not need to request discovery and
advertisements continually. Once a BTLE Peripheral (role that has
information) is advertised and a Central (role that receives
information) requests discovery, the app is allowed to go to
sleep. When a matching central or a peripheral device is found, iOS
wakes up the application app to handle such an event, even when the
application is in the background. MNA builds on this and leverages
asynchronous dispatch queues to turn each peer into a Central as well
as a Peripheral simultaneously. This architecture allows bidirectional
data transfers without concurrency issues.

@@

%
\section{Experimental Results}
\label{sec:eval}
%
Although MNA is deployed to about $10$ million users, collecting a
significant amount of data runs into privacy and scale issues. Also,
given the challenges associated with tests involving a large number of
physical devices along with human mobility patterns, we employ a
three-pronged approach to evaluation: (1) Limited metrics from the MNA
user population, (2) recruit human volunteers (colleagues at IBM) in a
pilot to have the app on their personal devices and run tests in the
background, and (3) use the automated test environment to run
controlled experiments. All experiments mentioning WiFi are on the
WiFi DNS channel and BT is Bluetooth classic.
%
\subsection{MNA limited experiment}
\label{sec:mna}
%
\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/mna}}
\caption{Messages received over mesh in one week}
\label{fig:mna}
\end{figure}
%
As Figure~\ref{fig:mna} shows, measures of number of messages received
first via MNA and the number of unique users was collected during a week in
June 2018. Almost no MNA messages were received with a roll-out to
$10\%$ of the users.  This makes sense given that p2p needs critical
density of users to propagate messages over large distances.  With
$100\%$ of the users participating, 6 million+ weather alerts were
received first via MNA by 1.9 million+ users within the collection
period. Even when the users are connected to the Internet, they
may receive the alert via MNA before the push notification arrives via
the Internet. Given that such a large population of users received it
via MNA first implies that there is value in the c2p2p architecture
even when Internet connectivity is available.
%
\subsection{Pilot test results}
\label{sec:romania}
%
About 25 colleagues participated in a pilot on a site across two
office buildings and about a thousand people in total.  Clearly, the
user density in this case was low. The app installed on their personal
devices sent a broadcast message of varying sizes every five minutes
and statistics on MNA were collected.

Overall, about 13\% of the messages were received at least once
(depends on density of users in the building) while users actively
used the devices for ~5\% of the time.  Figure~\ref{fig:romania_lat}
shows number of messages received and how the single-hop latency
varies between WiFi DNS and BT Classic and across increasing packet
sizes. Except for the 1000 byte+ packets, WiFi DNS achieved lower
end-to-end latencies than BT Classic, perhaps due to its broadcast
model and wide support across devices. This also explains why more
messages were delivered over WiFi DNS than BT Classic. Packet sizes
did not affect single-hop latencies on p2p DNS, perhaps because most
of the messages fit within a single advertised chunk of up to 750
bytes.

\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/romania_latency}}
\caption{Latency as a function of packet sizes}
\label{fig:romania_lat}
\end{figure}

%\begin{figure}[htbp]
%\centerline{\includegraphics[width=\columnwidth]{figs/romania_activity}}
%\caption{MNA activity during the pilot}
%\label{fig:romania_act}
%\end{figure}

%
\subsection{Automated test results}
\label{sec:automated}
%
We turn to the automated testing with up to 10 physical devices to
carry out some controlled testing and a deeper understanding of how
the various MNA protocols behave.
%
\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/variety_e2e_latency}}
\caption{End-to-end latency, WiFi DNS vs. BT Classic}
\label{fig:variety_e2e}
\end{figure}
%
Figure~\ref{fig:variety_e2e} shows end-to-end latency results in a test
involving three Android devices, each sending a variety of message and
payload size configurations. Latency is shown on a logarithmic scale
y-axis. Total load on network increases from left to right on
x-axis while the number of messages is small for the first and third
configurations and large for the second and fourth configuration.

As expected, WiFi DNS latency increases significantly with network
load given its very low throughput. As BT Classic cannot broadcast, it
spends significant time making connections with peers.  Hence, WiFi
DNS beats BT Classic in the 30x67 configuration involving a large
number of small messages. However, when the message sizes are
increased as in 42x117 configuration, BT Classic is almost 10x faster
than WiFi DNS due to superior
throughput. Figure~\ref{fig:variety_hops} confirms this difference in
network topology by measuring average hop counts. WiFi DNS always
needs fewer hops to deliver messages than BT Classic.
%
\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/variety_hops}}
\caption{Hop counts, WiFi DNS vs. BT Classic}
\label{fig:variety_hops}
\end{figure}
%

Next, we compare WiFi DNS and BT Nearby on Android with BTLE on iOS in
terms of end-to-end latency as the payload size is increased as shown
in Figure~\ref{fig:e2e}. The results were collected over three repeats
of the experiment involving up to 10 android devices and 6 iOS
devices. BTLE is almost 100x faster than WiFi DNS and BT Nearby. For
Wifi DNS, as payload size approaches the max advertisement chunk size
of 750 bytes, the end-to-end latency rises by almost 10x (100 bytes
versus 720 bytes). Behavior of BT Nearby taking time to form
connections but having a good throughput is evident as the end-to-end
latency does not change even when payload sizes go up to 10KB. BTLE on
the other hand gets 10x slower with 10KB payloads as compared to 1KB.

\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/e2e_latency}}
\caption{End-to-end latency, WiFi DNS vs. BT Nearby vs. BTLE}
\label{fig:e2e}
\end{figure}

As described above, MNA employs zero-overhead routing with the goal of
maximizing coverage and minimizing latency across peers while
tolerating topology changes.  As a result, messages may be delivered
multiple times to the same peer as no control messages are sent to
keep track of the peer state.  Figure~\ref{fig:dup} shows the cost of
such an approach in terms of average number of duplicate messages
received by peers across WiFi DNS, BT Nearby, and BTLE (iOS).

For small payload sizes, WiFi DNS ends up duplicating a large number
of messages as the broadcasts are seen by many peers and repeated
advertisements must go on to reach newly proximal peers. The
duplicates drop significantly with payloads approaching the
advertisement chunk size because the fraction of transmissions that
are successfull also drops with large payloads (multiple chunks need
to be delivered per message). The forward list in BT Nearby and BTLE
help control the duplicates significantly.

\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/duplicates}}
\caption{Duplicity due to zero-overhead routing}
\label{fig:dup}
\end{figure}
